% Chapter 2
\chapter{Literature Review}

This is the literature review chapter. Write your content here.

It should include a comprehensive overview of the existing research related to your topic. Discuss key findings, methodologies, and gaps in the literature that your thesis aims to address.
You can also include figures, tables, and citations to support your discussion. For example, you might reference a study that has significantly influenced your research direction.
Make sure to format your citations properly using the \texttt{natbib} package, as shown in the main thesis file. For instance, you can cite a source like this \citet{author2020study}
You can also include equations, if necessary, using the \texttt{amsmath} package. For example, you might want to present a mathematical model that is relevant to your research.
sample equation: \( E = mc^2 \) asd


\section{Key Findings}
Discuss the key findings from the literature. Highlight the most significant contributions and how they relate to your research question.
\section{Methodologies}
Describe the methodologies used in the studies you are reviewing. This could include experimental designs, data collection methods, and analytical techniques.
\section{Gaps in the Literature}
Identify the gaps in the existing literature that your research aims to fill. Discuss why these gaps are
important and how addressing them will contribute to the field.
\section{Conclusion}
Summarize the main points of your literature review. Emphasize the relevance of the reviewed